Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2017-12-26 survey/survey.tex
more work on intro

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reversibility, the Missing Principle}

@@

These quotes by Toffoli and Feynman both highlight two obvious
observations: (i) all the operations that a computer performs reduce
to basic physical operations; and (ii) there is a mismtach between the
logical operations of a typical model of computation (which are
logically irreversible) and the fundamental laws of physics (which are
reversible). One could certainly dismiss the mismatch but our thesis
is that there is great value in revisiting models of computation to be
in harmony with the laws of physics.

@@

\subsection{Maxwell's Demon}

To fully appreciate one of the principles is missing in conventional computing,
one must go back to an old thought experiment by J. C. Maxwell. This is
codified in a letter that Maxwell wrote to P. G Tait in 1867 --
the letter whose ideas are now known as `Maxwell's
Demon'. Maxwell's Demon is a thought experiment that seems to
indicate that intelligent beings can somehow violate the second law of
thermodynamics, thereby violating physics itself.

Many resolutions were offered for this conundrum (for a compilation, see
Maxwell's Demon~\cite{leff1990}), but none withstood careful
scrutiny until the establishment of `Landauer's Principle' in 1961 -- a
principle whose experimental validation
happened recently in 2012~\cite{berut2012experimental}.

@@

This is a strange and wonderful idea: Information,
physics and computation are inextricably linked. Furthermore, this
implies that ideas in each field have consequences for the other.
(Cite Bennett and the various `thermodynamics of computation' papers here.)

@@

\begin{quote}
Toffoli 1980: Mathematical models of computation are abstract
constructions, by their nature unfettered by physical laws. However,
if these models are to give indications that are relevant to concrete
computing, they must somehow capture, albeit in a selective and
stylized way, certain general physical restrictions to which all
concrete computing processes are subjected.
\end{quote}

@@

\subsection{Origins}

@@

\subsection{From Classical to Quantum}

@@

\subsection{Conservation of Information}

@@

\subsection{Information}

@@

%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Programs as Reversible Deformations}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data I: Finite Sets}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data II: Structured Finite Types}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data III: Reversible Programs between Reversible Programs}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements} We would like to thank the numerous
students and colleagues who participated in various aspects of this
research and who provided valuable feedack and constructive criticism.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Macros

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reversibility, the Missing Principle}

@@

These quotes by Toffoli and Feynman both highlight two obvious
observations: (i) all the operations that a computer performs reduce
to basic physical operations; and (ii) there is a mismtach between the
logical operations of a typical model of computation (which are
logically irreversible) and the fundamental laws of physics (which are
reversible). One could certainly dismiss the mismatch as irrelevant to
the practice of computing but our thesis is that there is the next
computing revolution is likely to be founded on revised models of
computation that are designed to in harmony with the laws of physics.

After a detailed introduction on the origins of ``logically
reversibile computer operations'' and an excursion in the origins of
our the status quo based on ``irreversible computer operations'', we
will develop in detail three reversible models of computation and
discuss their potential applications.

@@

%%%%%%%%%
\subsection{Maxwell's Demon}

To fully appreciate one of the principles is missing in conventional
computing, one must go back to an old thought experiment by
J. C. Maxwell. This is codified in a letter that Maxwell wrote to P. G
Tait in 1867 -- the letter, whose ideas are now known as
\emph{Maxwell's Demon}, tells of a thought experiment that seems to
indicate that intelligent beings can somehow violate the second law of
thermodynamics, thereby violating physics itself. Many resolutions
were offered for this conundrum (for a compilation, see Maxwell's
Demon~\cite{leff1990}), but none withstood careful scrutiny until the
establishment of \emph{Landauer's Principle} in 1961 -- a principle
whose experimental validation happened recently in
2012~\cite{berut2012experimental}.

@@

This is a strange and wonderful idea: information, physics and
computation are inextricably linked. Furthermore, this implies that
ideas in each field have consequences for the
other~\cite{bennett:1973:lrc,bennett1985fundamental,bennett2010notes,bennett2003notes,baker:1992:nft,dblp:conf/csfw/malacarias12}.
%% (Cite Bennett and the various `thermodynamics of computation'
%% papers here.)

@@

@@

%%%%%%%%%
\subsection{Origins of Computational Models}

@@

%%%%%%%%%
\subsection{From Classical to Quantum}

@@

%%%%%%%%%
\subsection{Conservation of Information}

@@

%%%%%%%%%
\subsection{Information}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Programs as Reversible Deformations}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data I: Finite Sets}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data II: Structured Finite Types}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data III: Reversible Programs between Reversible Programs}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements} We would like to thank the numerous
students and colleagues who participated in various aspects of this
research and who provided valuable feedack and constructive criticism.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
